An Exposition on the Propriety of Restricted Boltzmann Machines

### 2016-07-28 presentations/jsm2016/index.Rmd
slides complete, but check densities

## A tale of three priors

>>>>

## A tale of three methods

> 1. *Trick prior.* Cancel out the normalizing term, full conditionals normally dist'd. 
   $$
    \pi(\theta) \propto \gamma(\theta)^n \exp\left(-\frac{1}{2C_{1}} \theta_{main}' \theta_{main} -\frac{1}{2C_{2}} \theta_{int}' \theta_{int}\right),
   $$
    where $\gamma(\theta)$ is the normalizing term [@li2014biclustering].
> 2. *Truncated Normal prior.* Use two independent truncated spherical normal distributions as priors for $\theta_{main}$ and $\theta_{int}$ with $\sigma_{int} < \sigma_{main}$. Not conjugate $\Rightarrow$ geometric adaptive MH step [@zhou2014some] and calculation of $\gamma(\theta)$. 
> 3. *Marginalized likelihood* Marginalize out $h$ in $f_{\theta}(x)$, use trunc. Normal prior. 
    $$
    g_{ \theta}(v) = \sum\limits_{\in \{-1,1\}^H} \exp\left(\sum\limits_{i = 1}^V \sum\limits_{j=1}^H \theta_{ij} v_i h_j + \sum\limits_{i = 1}^V\theta_{v_i} v_i + \sum\limits_{j = 1}^H\theta_{h_j} h_j\right)
    $$

## Posterior distributions of images

```{r models, cache=TRUE}
source("../../4x4 model/functs_sample.R")

#data and params ------------------------
H <- 4
V <- 4

#marginalized likelihood
load("../../4x4 model/written/fitted_models_trunc_marginal_thin.Rdata")
marginal_bad <- models_bad
marginal_good <- models_good

#load trick prior
load("../../4x4 model/written/fitted_models_jing_5.8.Rdata")
trick_bad <- models_bad
trick_good <- models_good

#truncated normal
load("../../4x4 model/written/fitted_models_trunc_thin.Rdata")
trunc_bad <- models_bad
trunc_good <- models_good

#rm unneccesary data
rm(models_bad)
rm(models_degen)
rm(models_good)

load("../../4x4 model/written/params_theta.Rdata")

params_good <- list(main_hidden = sample.params %>% ungroup() %>% filter(!near_hull) %>% select(starts_with("h"), -H) %>% data.matrix(),
                    main_visible = sample.params %>% ungroup() %>% filter(!near_hull) %>% select(starts_with("v"), -V) %>% data.matrix(),
                    interaction = sample.params %>% ungroup() %>% filter(!near_hull) %>% select(starts_with("theta")) %>% data.matrix() %>% matrix(4))

#computing actual distributions ---------------------
distn_good <- visible_distn(params = params_good)

reshape_sample_distn <- function(model) {
  sample_distn <- model$distn
  dim(sample_distn) <- c(dim(sample_distn)[1]*dim(sample_distn)[2], dim(sample_distn)[3])
  sample_distn %>% data.frame() -> sample_distn
  names(sample_distn) <- names(distn_good)

  sample_distn %>%
    group_by(image_id) %>%
    mutate(iter = 1:n()) -> sample_distn

  return(sample_distn)
}

marginal_sample_good <- reshape_sample_distn(marginal_good)
trick_sample_good <- reshape_sample_distn(trick_good)
trunc_sample_good <- reshape_sample_distn(trunc_good)

marginal_sample_good %>% rename(marginal = prob) %>%
  left_join(trick_sample_good %>% ungroup() %>% select(-image_id) %>% rename(trick = prob)) %>%
  left_join(trunc_sample_good %>% ungroup() %>% select(-image_id) %>% rename(trunc = prob)) %>%
  ungroup() %>%
  mutate(image_id = paste0("image_", image_id)) %>% 
  gather(method, prob, trick, marginal, trunc) %>%
  spread(image_id, prob) -> all_statistics

all_statistics %>%
  gather(statistic, value, -iter, -method, -starts_with("v")) %>%
  filter(grepl("image", statistic) & !is.na(value)) %>%
  separate(statistic, into = c("junk", "statistic")) %>%
  mutate(statistic = factor(statistic, labels = paste("Image", 1:length(unique(statistic))))) %>%
  left_join(distn_good %>% select(-image_id)) %>% 
  ggplot() +
  geom_density(aes(value, colour = method, fill = method), alpha = .2) +
  geom_vline(aes(xintercept = prob)) +
  facet_wrap(~statistic, scales = "free_x") + 
  ylab("Posterior Density") + xlab("Probability of Image") +
  scale_colour_discrete("Method", labels = c("Marginalized likelihood", "Trick prior", "Truncated Normal prior")) +
  scale_fill_discrete("Method", labels = c("Marginalized likelihood", "Trick prior", "Truncated Normal prior")) -> p.models
```
```{r models-plot, fig.width=8, fig.height=4}
p.models
```
Each of these rigorous fitting methods seeks to merely replicate the empirical data distribution, which is the optimal nonparametric solution.
